Survey of Cache Oblivious B-trees

### 2011-12-11 paper.tex
added some figures for string b-tree section

\section{Variable Length Strings}

\subsection{Motivation}
In order for B-trees to remain versatile, they must be able to accommodate various forms of data. However, traditional B-trees and string B-trees have suboptimal performance when variable length strings are used as keys. In addition, string B-trees do not support compression, and commonly used front compression is handled sub-optimally. Regular B-trees and string B-trees are not cache oblivious, and thus incur poorer performance as described previously. Range queries have poor performance in B-trees. In \cite{BenderFaKu06}, Bender et al. address these issues by designing a Cache Oblivious String B-tree (COSB-tree), which they argue works efficiently in these problem areas for variable length string keys. They also provide a modified front-compression scheme, which gives their approach the advantages of using less space and having improved locality within the compressed data.

@@

\subsubsection{The Dictionary Matching Problem and Centroid Trees} %todo:cite dictionary matching problem--sources in cosb-tree paper
In this approach to solving the dictionary matching problem *cite dictionary matching problem paper*, a dictionary $D$ of $N$ keys is preprocessed using hashing and divide-and-conquer via the construction of a centroid tree over the keys. This technique forms the basis for the Cache Oblivious String B-tree (COSB-tree) design. Consider the compacted trie $T$ of $D$. Bender et al. observe that there exists a centroid vertex $\rho$ in $T$ such that $\rho$ has between N/3 and 2N/3 descendants, inclusive. When searching for a key $k$ in the compacted trie, you traverse the tree as follows: for each node $y$ that you visit, if $y$ is a prefix of $x$, or $y$ matches $x$, continue on into the trie rooted at $y$ or \emph{down trie} of $y$. Otherwise traverse the trie resulting from excluding $y$ and its subtree. The \emph{centroid tree} is obtained by rooting the tree at \emph{centroid vertex} $\rho$, and making its children recursively defined centroid trees of its up and down tries.

>>>>

\section{Variable Length Strings}
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/COSB-treegeneralstructurediagram.pdf}
\end{center}

\caption{An illustration of the dynamic COSB-tree structure}
\label{fig:string_structure}
\end{figure}
\subsection{Motivation}
In order for B-trees to remain versatile, they must be able to accommodate various forms of data. However, traditional B-trees and string B-trees have suboptimal performance when variable length strings are used as keys. In addition, string B-trees do not support compression, and commonly used front compression is handled sub-optimally. Regular B-trees and string B-trees are not cache oblivious, and thus incur poorer performance as described previously. Range queries have poor performance in B-trees. In \cite{BenderFaKu06}, Bender et al. address these issues by designing a Cache Oblivious String B-tree (COSB-tree), which they argue works efficiently in these problem areas for variable length string keys. They also provide a modified front-compression scheme, which gives their approach the advantages of using less space and having improved locality within the compressed data.

@@

\subsubsection{The Dictionary Matching Problem and Centroid Trees} %todo:cite dictionary matching problem--sources in cosb-tree paper
\begin{figure}

\begin{center}
	\includegraphics[width=0.8\columnwidth]{figures/centroidtreeexample.pdf}
\end{center}

\caption{A centroid tree example}
\label{fig:centroid}
\end{figure}
In this approach to solving the dictionary matching problem *cite dictionary matching problem paper*, a dictionary $D$ of $N$ keys is preprocessed using hashing and divide-and-conquer via the construction of a centroid tree over the keys. This technique forms the basis for the Cache Oblivious String B-tree (COSB-tree) design. Consider the compacted trie $T$ of $D$. Bender et al. observe that there exists a centroid vertex $\rho$ in $T$ such that $\rho$ has between N/3 and 2N/3 descendants, inclusive. When searching for a key $k$ in the compacted trie, you traverse the tree as follows: for each node $y$ that you visit, if $y$ is a prefix of $x$, or $y$ matches $x$, continue on into the trie rooted at $y$ or \emph{down trie} of $y$. Otherwise traverse the trie resulting from excluding $y$ and its subtree. The \emph{centroid tree} is obtained by rooting the tree at \emph{centroid vertex} $\rho$, and making its children recursively defined centroid trees of its up and down tries.
